Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-16 WSSSPE1_report/WSSSPE1_summary_report.tex
Section 3.1: Keynote 1:  Made changes the speaker requested.
text improvement

The first keynote~\cite{WSSSPE1-keynote1} was delivered by Philip
E. Bourne of University of California, San Diego.  Bourne is a basic
biomedical scientist but has formed four software companies. He also
helped found PLOS~\cite{plos-web} and PDB~\cite{pdb-web}.  He is
working on automating three-dimensional visualizations of cell
contents and molecular structures, a problem that has not been solved
and when done, would serve as a key function of software in biomedical
sciences.

@@

The Christopherson paper reports on the application of Steps 1-3 to a computational modeling framework developed in the 1990s.
Step 1 was realized as a 2-day, in person specifications meeting and code walk-through. Step 2 was a 5-day hack-athon \todo{we spell hackathon three different ways in two paragraphs - Pick one and be consistent.} to develop working code.
Step 3 was a 3-day hack-a-thon to refine the code based on new requirements. The team worked on small, low-risk unit of
code. It was challenging, revealed unanticipated obstacles, programmers had to work together, and experimentation was encouraged.

@@

\section{Illustrative Use Cases} \label{sec:use-cases}

\note{lead: Ketan Maheshwari, contributors: Andy Terrel}

Following is a list of software projects discussed throughout the workshop:

\begin{itemize}
    \item mvapich, Cactus, DUNE
    \item gateways, github, VisIT, visTrails, PAAS
    \item SimGrid, Swift, Galaxy
    \item FastMath, R/qtl, MINRES-QLP
    \item OWL/RDF, AMR/multiphysics, Globus-Genomics
\end{itemize}

In this section, we discuss the following:
\begin{itemize}
    \item Categorize the real-world software projects into workshop themes. This is based on the aspect of a software project that dominates and best fits one of the workshop themes.
    \item How the worshop themes are applicable to and has impacted software projects over its lifetime.
    \item Similarities and differences between these projects.
    \item What is a takeaway for a new project? What are the lessons learned? How can the course of current and future development projects be altered to implement lessons learned.
\end{itemize}

The software projects discussed in the workshop can be classified as follows:
\begin{itemize}
\item General Purpose Software applicable to a generic task and/or enabling other software to run by providing suitable environment. eg. Collaborative development, Workflow and scripting frameworks, visualization frameworks.
\item Special Purpose Software that was developed specifically for a target use-case scenario or keeping some fixed/frozen requirements in mind.
\begin{itemize}
  \item Domain science oriented software developed to solve domain science problems, eg. MINRES-QLP.
  \item Infrastructure oriented software developed to better utilize new infrastructure, eg. MVAPICH.
\end{itemize}
\end{itemize}

A recurrent theme of the workshop was following best practices while developing software. Many submissions discuss software development best practices.

>>>>

The first keynote~\cite{WSSSPE1-keynote1} was delivered by Philip
E. Bourne of University of California, San Diego.  Bourne is a basic
biomedical scientist but has formed four software companies. He co-founded 
PLOS Computational Biology~\cite{plos-web} and helped develop the RCSB PDB~\cite{pdb-web}.  
He is working on automating three-dimensional visualizations of cell
contents and molecular structures, a problem that has not been solved
and when done, would serve as a key function of software in biomedical
sciences.

@@

The Christopherson paper reports on the application of Steps 1-3 to a computational modeling framework developed in the 1990s.
Step 1 was realized as a 2-day, in person specifications meeting and code walk-through. Step 2 was a 5-day hackathon to develop working code.
Step 3 was a 3-day hackathon to refine the code based on new requirements. The team worked on small, low-risk unit of
code. It was challenging, revealed unanticipated obstacles, programmers had to work together, and experimentation was encouraged.

@@

\section{Case Studies} \label{sec:use-cases}
\note{lead: Ketan Maheshwari, contributors: Andy Terrel}

For the purpose of studying software case-studies we classify the software
projects discussed in the workshop into three broad categories. First, the
`utility software', software tools that enable and/or facilitate the
development of other tools and techniques to carry out scientific work. Second,
the `scientific software', software that was developed with an aim to solve a
class of scientific problems. Third, the `infrastructure software', software
developed to efficiently utilize new research infrastructure.

Although we anticipate overlap in categories and point it out in our
discussion. The classification is based on two criterions: First, the character
of a software project that dominates and best fits and the theme the paper
discussing the project was placed in the workshop discussion.

We discuss the case studies with an aim of understanding the practicalities of
the many points discussed in the workshop and hope to extract the best
practices and associate them with the projects. We discuss how the cases are
applicable to and has impacted user communities over their lifetime isolating
similarities and differences between projects. What are the takeaways for a new
project and the lessons learned?  How can the course of current and
future development projects be altered to implement lessons learned.

\subsection{Utility Software}
The General Purpose Software applicable to a generic task and/or enabling other
software to run by providing suitable environment. For example, collaborative
development, Workflow and scripting frameworks, visualization frameworks.
\subsection{Scientific Software}
Special Purpose Software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind.  
Domain science
oriented software developed to solve domain science problems, eg MINRES-QLP.
\subsection{Infrastructure Software}
Infrastructure oriented software developed to better utilize new
infrastructure, eg. MVAPICH.
%
%\begin{itemize}
%  \item mvapich, Cactus, DUNE
%  \item gateways, github, VisIT, visTrails, PAAS
%  \item SimGrid, Swift, Galaxy
%  \item FastMath, R/qtl, MINRES-QLP
%  \item OWL/RDF, AMR/multiphysics, Globus-Genomics
%\end{itemize}
%
